```{r echo = FALSE}
knitr::opts_chunk$set(out.width = "100%")
```

# (PART\*) Create Workflows With Galaxy {-}   

# Overview {#create-workflows-with-galaxy-overview}

## Learning Objectives

1. Learn how to launch Galaxy in AnVIL and import data
1. Understand how to find, import, and run workflows
1. Visualize a sample variant using JBrowse
1. Learn how to save your history back to AnVIL
1. Show how to shut down Galaxy in AnVIL

# Preparation {#create-workflows-with-galaxy-preparation}

## Review Key Concepts

This 5-min video provides a high level summary of the exercises to follow.  Several important concepts are introduced to provide context for the exercises ([slides](https://docs.google.com/presentation/d/1nZIlyA-YIxRDYUiozl1UTXIhJJeBb7VbqlXWig7jXAk)).

## Clone Workspace

:::{.notice}
For this demo, you will need your own copy of the [`SARS-CoV-2-Genome`](https://anvil.terra.bio/#workspaces/gdscn-exercises/SARS-CoV-2-Genome) Workspace.  These instructions will walk you through cloning a copy of the Workspace.
:::

## Start Cloud Environment

:::: {.borrowed_chunk}
```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_galaxy_start.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```
::::

# Exercises {#create-workflows-with-galaxy-exercises}

## Import Data

```{r, echo=FALSE}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1KnuhIqChw7dD6IuG8T35bfqMRpvkraQGkkc7mDhHbpk/edit#slide=id.g25530dc4fb7_0_0" )
```

## Find Workflow

```{r, echo=FALSE}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1KnuhIqChw7dD6IuG8T35bfqMRpvkraQGkkc7mDhHbpk/edit#slide=id.g25530dc4fb7_0_149" )
```

## Run Workflow

```{r, echo=FALSE}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1KnuhIqChw7dD6IuG8T35bfqMRpvkraQGkkc7mDhHbpk/edit#slide=id.g25530dc4fb7_0_298" )
```

## Explore Results

```{r, echo=FALSE}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1KnuhIqChw7dD6IuG8T35bfqMRpvkraQGkkc7mDhHbpk/edit#slide=id.g25530dc4fb7_0_447" )
```

## Export History

```{r, echo=FALSE}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1KnuhIqChw7dD6IuG8T35bfqMRpvkraQGkkc7mDhHbpk/edit#slide=id.g25530dc4fb7_0_599" )
```

## Shut Down

```{r, echo=FALSE}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1KnuhIqChw7dD6IuG8T35bfqMRpvkraQGkkc7mDhHbpk/edit#slide=id.g25530dc4fb7_0_748" )
```

# Instructor Guide {#create-workflows-with-galaxy-instructor-guide}

## Timeline

