```{r echo = FALSE}
knitr::opts_chunk$set(out.width = "100%")
```

# (PART\*) Human Genetic Variation in MAGE {-}   

# Overview {#human-genetic-variation-in-mage-overview}

Introduce AnVIL, Bioconductor, and Galaxy using MAGE

## Learning Objectives

1. Explore MAGE Workspace 
1. Analysis with Jupyter/Terminal
1. Bioconductor with RStudio
1. Workflows with Galaxy

# Preparation {#human-genetic-variation-in-mage-preparation}

## Review Background

## Create AnVIL Account

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_0")
```

## Clone Workspace

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g36877bec10a_0_0")
```

# Exercises {#human-genetic-variation-in-mage-exercises}

## Explore MAGE Workspace

### Review Key Concepts

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g33d1f69b524_1_1")
```

### Scavenger Hunt

Original Workspace

- https://anvil.terra.bio/#workspaces/anvil-outreach/demos-explore-mage
- Explore Data Tables e.g. counts.csv, PCA.nb.html
- Find files in Workspace Bucket

Your Cloned Workspace

- https://anvil.terra.bio/#workspaces/ &lt;billing-project&gt; / &lt;workspace-name&gt;
- Explore Data Tables
- Look for files in Workspace Bucket

## Analysis with Jupyter/Terminal

### Launch Jupyter

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1-IU5l40UxmpoBqFimSUmT1PNEUij2UnXMslOsiG5ncE/edit#slide=id.g14ea2db115d_0_22")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1-IU5l40UxmpoBqFimSUmT1PNEUij2UnXMslOsiG5ncE/edit#slide=id.g14ea2db115d_0_18")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1-IU5l40UxmpoBqFimSUmT1PNEUij2UnXMslOsiG5ncE/edit#slide=id.g14ea2db115d_0_41")
```

### Review Key Concepts

### Enter Terminal

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_32")
```

### Run Jupyter Notebook

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_36")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1-IU5l40UxmpoBqFimSUmT1PNEUij2UnXMslOsiG5ncE/edit#slide=id.g25442045568_0_82")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_40")
```

### Next Steps

## Bioconductor with RStudio

### Launch RStudio

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1a35Mb8f0M-bQkBcHa1cyQc6YxXoBLtExCz96nv08vkA/edit#slide=id.g14ea2db115d_0_18")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1a35Mb8f0M-bQkBcHa1cyQc6YxXoBLtExCz96nv08vkA/edit#slide=id.g256428d32e5_0_10")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_71")
```

### Review Key Concepts

### Use AnVILGCP

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_80")
```

### Run R Notebook

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_84")
```

### Next Steps

## Workflows with Galaxy

### Launch Galaxy

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1yYCg4cPVBMMDghT17B4XzROieqyMH99Ex9nMm_Scm9Q/edit#slide=id.ge1182914a6_0_36")
```

### Review Key Concepts

### Run FastQC

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_90")
```

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_96")
```

### Install Tool

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_100")
```

### Next Steps

# Wrap Up {#human-genetic-variation-in-mage-wrap-up}

## Shut Down Compute

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_110")
```

## Learn More

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_126")
```

## Provide Feedback

```{r, echo=FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/16DHXD2KNyjjP2mKzPDHmNE9loYy6OgqGG0-Kn8CLFak/edit#slide=id.g369240085c8_0_120")
```
